Relative parabolic factorizations of Chevalley groups

### 2015-05-01 bass-kolster.tex
BK for A_l, C_l fixes

\noindent Then the principal congruence subgroup $\G(\Phi,R,I)$ admits the following relative version of Bass---Kolster decomposition:
\[\G(\Phi,R,I)=\U(\Sigma_1^-\setminus\{-\alpha_\mathrm{max}\},I)\cdot \U(\Sigma_1^+,I)\cdot \Z\cdot \U(\Sigma_1^-,I)\cdot \U(\Sigma_1^+,I)\cdot \G(\Delta_1,R,I), \]
where $\Z=\left\{ z_{-\alpha_\mathrm{max}}(r,1)\mid r\in I \right\}$.
\end{thm}
\begin{proof}
Let $g$ be an element of $\G(\Phi, R, I)$. It is clear that vector $v=g\cdot v^+$ is $I$-unimodular.
In view of relative Chevalley---Matsumoto lemma it suffices to show that $\Z \cdot \U(\Phi^-, I) \cdot \U(\Phi^+, I) \cdot g$ contains an element $g$ whose matrix entry at position $(1,1)$ is equal to $1$.
We will need to analyze several cases.

\textsc{Case $\Phi=\rA_\ell$:}
Denote $v=(1+r_1,r_2,\ldots,r_\ell,r_{\ell+1})^t\in\Ums(\ell+1,R,I)$.
One can add the last entry to the first $\ell$ with some coefficients from $I$ to make them form a unimodular column of length $\ell$.
Then $\ell$ lower transformations (with coefficients from $I$) can make the last entry equal $r_1$.
Now apply $z_\gamma(-r_1,1)$, $\gamma=-\alpha_{\max}$:
\begin{multline*}
(1+r_1,r_2,\ldots,r_\ell,r_1)^t\xmapsto{x_{-\gamma}(-1)}(1,r_2,\ldots,r_\ell,r_1)^t\xmapsto{x_\gamma(-r_1)}\\\longmapsto(1,r_2,\ldots,r_\ell,0)^t\xmapsto{x_{-\gamma}(1)}(1,r_2,\ldots,r_\ell,0)^t.
\end{multline*}

@@

Now we can assume that the first $2\ell-1$ entries of $v$ are unimodular and find $c_1,\ldots,c_{-2}\in I$ such that $c_1v_1+\ldots+c_{-2}v_{-2}=v_1-v_{-1}$. Add $c_{-i}v_i$ to $v_{-1}$, $i=2,\ldots,\ell$:
\[ (v_1,v_2\ldots,v_\ell,v_{-\ell},\ldots,v_{-2},v_{-1})^t\longmapsto (v_1,v'_2\ldots,v'_\ell,v_{-\ell},\ldots,v_{-2},v'_{-1})^t. \]
Then add $c_iv_i'=c_i(v_1+c_{-i}v_{-i})$, $i=2,\ldots,\ell$ to the last entry:
\[ v_1\longmapsto v_1,\quad v'_{-1}\longmapsto v_{-1}+\sum_{i=2}^\ell c_{-i}v_{-i}+\sum_{i=2}^\ell c_i(v_i+c_{-i}v_1)=v''_{-1}. \]
Next add $\left(c_1-\sum_{i=2}^\ell c_ic_{-i}\right)v_1$ to $v''_{-1}$ to get $v_1$ in position $-1$.
Again, as in case of $\rA_n$, apply $z_\gamma(-v_1,1)$ with $\gamma=-\alpha_\mathrm{max}$, to get $1$ as the first entry and $0$ as the last.

\textsc{Case $\Phi=\rD_\ell$.} 
By Lemma~\ref{lemma:asrUnip} we can find $h_1\in \U(\Sigma^+_\ell, I)$ such that the upper half $v'^+$ of vector $v'=h_1\cdot v$ is $I$-unimodular.
Since $\sr(I)\leq \ell-1$ we can find $c_1$, $c_3, \ldots c_\ell \in I$ such that $(v''_1, v''_3, \ldots, v''_\ell) \in \Ums(\ell-1, R, I)$, where
$v''=h_2 \cdot t_{1,2}(c_1) \cdot v'$, $h_2=\prod\limits_{i=3}^\ell t_{i,2}(c_i)$.

@@

\begin{rem} %TODO:
\textbf{TODO: Calculate the exact number of $z_\alpha$'s involved in the decomposition}

\textsc{$\Phi=\rA_\ell$}
Now $\ell-1$ lower transformations eliminates all the remaining entries in the first column, while $\ell$ more upper ones will make the first row trivial.

\textsc{$\Phi=\rC_\ell$}
Now apply $2\ell-2$ lower transformation to obtain zeroes on positions $i=2,\ldots,-2$, that will also add $\sum_{i=2}^\ell (v_i+v_{-i})$ to $v_{-1}$, 
so by one more transformation one can make the first column trivial. 
The same can be done with the first row, so in total we used $8\ell-3$ transformations.

\end{rem}

>>>>

\noindent Then the principal congruence subgroup $\G(\Phi,R,I)$ admits the following relative version of Bass---Kolster decomposition:
\[\G(\Phi,R,I)=\U(\Sigma_1^-\setminus\{-\alpha_\mathrm{max}\},I)\cdot \U(\Sigma_1^+,I)\cdot \Z\cdot \U(\Sigma_1^-,I)\cdot \U(\Sigma_1^+,I)\cdot \G(\Delta_1,R,I), \]
where $\Z=\left\{ z_{-\alpha_\mathrm{max}}(r,1)\mid r\in I \right\}$.
\end{thm}
\begin{proof}
Let $g$ be an element of $\G(\Phi, R, I)$. It is clear that vector $v=g\cdot v^+$ is $I$-unimodular.
In view of relative Chevalley---Matsumoto decomposition it suffices to show that $\Z \cdot \U(\Phi^-, I) \cdot \U(\Phi^+, I) \cdot g$ contains an element whose matrix entry at position $(1,1)$ is equal to $1$.
We will need to analyze several cases.

\textsc{Case $\Phi=\rA_\ell$:}
Denote $v=(1+v_1,v_2,\ldots,v_\ell,v_{\ell+1})^t\in\Ums(\ell+1,R,I)$.
One can add the last entry to the first $\ell$ with some coefficients from $I$ to make them form a unimodular column of length $\ell$.
Then $\ell$ lower transformations (with coefficients from $I$) can make the last entry equal to $v_1$.
Now apply $z_\gamma(-v_1,1)$, $\gamma=-\alpha_{\max}$:
\begin{multline*}
(1+v_1,v_2,\ldots,v_\ell,v_1)^t\xmapsto{x_{-\gamma}(-1)}(1,v_2,\ldots,v_\ell,v_1)^t\xmapsto{x_\gamma(-v_1)}\\\longmapsto(1,v_2,\ldots,v_\ell,0)^t\xmapsto{x_{-\gamma}(1)}(1,v_2,\ldots,v_\ell,0)^t.
\end{multline*}

@@

Now we can assume that the first $2\ell-1$ entries of $v$ are unimodular and find $c_1,\ldots,c_{-2}\in I$ such that $c_1v_1+\ldots+c_{-2}v_{-2}=(v_1-1)-v_{-1}$. Add $c_{-i}v_i$ to $v_{-1}$, $i=2,\ldots,\ell$:
\[ (v_1,v_2\ldots,v_\ell,v_{-\ell},\ldots,v_{-2},v_{-1})^t\longmapsto (v_1,v'_2\ldots,v'_\ell,v_{-\ell},\ldots,v_{-2},v'_{-1})^t. \]
Then add $c_iv_i'=c_i(v_1+c_{-i}v_{-i})$, $i=2,\ldots,\ell$ to the last entry:
\[ v_1\longmapsto v_1,\quad v'_{-1}\longmapsto v_{-1}+\sum_{i=2}^\ell c_{-i}v_{-i}+\sum_{i=2}^\ell c_i(v_i+c_{-i}v_1)=v''_{-1}. \]
Next add $\left(c_1-\sum_{i=2}^\ell c_ic_{-i}\right)v_1$ to $v''_{-1}$ to get $v_1-1$ in position $-1$.
Again, as in case of $\rA_n$, apply $z_\gamma(1-v_1,1)$ with $\gamma=-\alpha_\mathrm{max}$, to get $1$ as the first entry and $0$ as the last.

\textsc{Case $\Phi=\rD_\ell$.} 
By Lemma~\ref{lemma:asrUnip} we can find $h_1\in \U(\Sigma^+_\ell, I)$ such that the upper half $v'^+$ of the vector $v'=h_1\cdot v$ is $I$-unimodular.
Since $\sr(I)\leq \ell-1$ we can find $c_1$, $c_3, \ldots c_\ell \in I$ such that $(v''_1, v''_3, \ldots, v''_\ell) \in \Ums(\ell-1, R, I)$, where
$v''=h_2 \cdot t_{1,2}(c_1) \cdot v'$, $h_2=\prod\limits_{i=3}^\ell t_{i,2}(c_i)$.

@@

\begin{rem} %TODO:
\textbf{TODO: Calculate the exact number of $z_\alpha$'s involved in the decomposition}
\end{rem}
